---
title: "NE WNV Example"
author: "Kelly Helm Smith & Drew Tyre"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Functional Linear Modeling for Vector-borne Disease}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up data for example
```{r data, message=FALSE}
library(tidyverse)
library(flm.NE.WNV) 

# Read in population data
pop = flm.NE.WNV::NE_county_pops  %>% 
  filter(year >= 2002) %>% 
  as.data.frame()
 
county_lookup <- pop %>% 
  filter(year == "2010") %>% 
  select(County, fips)%>% 
  as.data.frame()

#Human cases, simulated, based on predicted values from the fitted model. 
cases = flm.NE.WNV::sampledat %>% 
  select(County, year, cases)%>% 
  # # drop Arthur back to 0 to test new zero filling methods
  # mutate(cases = case_when(County == "Arthur" ~ 0,
  #                          TRUE ~ cases)) %>% 
  as.data.frame()

# Read in environmental data
NEdat = flm.NE.WNV::NEdat %>% 
  rename(tmean = temp) %>% 
  mutate(fips = sprintf("31%03d", cofips)) %>% 
  left_join(county_lookup, by = "fips") %>% 
  select(County, fips, year, month, tmean, ppt)%>% 
  as.data.frame()
  

# Read in spi
spi = flm.NE.WNV::spi %>% 
  mutate(County = trimws(County),
         County = gsub(" ", "", County)) %>% 
  left_join(county_lookup, by = "County") %>% 
  rename(spi = spi1) %>% 
  select(County, fips, year, month, spi)%>% 
  as.data.frame()

# Read in spei
spei = flm.NE.WNV::spei %>% 
  mutate(County = trimws(County),
         County = gsub(" ", "", County)) %>% 
  left_join(county_lookup, by = "County") %>% 
  rename(spei = spei1) %>% 
  select(County, fips, year, month, spei)%>% 
  as.data.frame()

# Specify the last date to use -- package data incomplete for 2019
target.date = "2012-02-01" 
#target.date = "2019-03-01" -- this should fail with an error
# Specify the first year of data to use
start.year = 2002

# Specify lag lengths to use
lag.lengths = c(12, 18) #c(12, 18, 24, 30, 36)

#For consistent results with simulated data
in.seed = 4872957 # set = NULL if you have actual case data

```
## Some basic quality control on synthetic data

Our approach struggles if there are counties with no cases, so we eliminate those counties (two in the simulated data).

```{r}
data <- assemble.data.lags(pop, cases, NEdat, spi, spei, 
                           target.date, start.year, in.seed, lag.lengths)
allLagsT <- data[[1]] # 2004 - 2017
allLags0 <- data[[2]] # 2018
# how many counties lack cases? Should be none.
allLagsT %>% group_by(County) %>% 
  summarize(total_cases = sum(cases)) %>% 
  filter(total_cases == 0)

allLagsT %>% group_by(County) %>% 
  summarize(total_cases = sum(cases)) %>% 
  filter(total_cases > 0)

# to identify counties to fill in
allunits <- unique(cases$County)
missingunits <- !(allunits %in% unique(allLagsT$County))
allunits[missingunits]
```

What does the state total look like plotted against time?

```{r}
allLagsT %>% group_by(year) %>% 
  summarize(total_cases = sum(cases)) %>% 
  ggplot() + 
  geom_point(mapping = aes(x = year, y = total_cases))
```

## Run the model

This chunk assembles data and runs models, using all combinations of lagged data for mean temperature, precipitation, SPI and SPEI.

Note: It takes about 5 minutes to run the models with a lag of 9. It takes 25 minutes to run all the models with the set of default lags.


```{r flm}

flm.results = call.flm(pop, cases, NEdat, 
                       spi, spei, target.date, start.year, 
                       in.seed = in.seed, lag.lengths,
                       fillzeros = TRUE,
                       nsim = 10)

```

## Overview of model chosen by AIC 
```{r}
top <- flm.results[[3]]$best

mod <- summary(flm.results[[3]]$fittedModels[[top]])
mod

```
## View predictions from best-fit model
```{r}
flm.results[[1]]
```

```{r}
dist_lag_terms <- extract_functional(flm.results[[3]]$fittedModels[[top]]) %>% 
  mutate(lcl = fit - 1.96*se,
         ucl = fit + 1.96*se)
ggplot(data = dist_lag_terms) + 
  geom_line(mapping = aes(x = x, y = fit)) +
  geom_ribbon(mapping = aes(x = x, ymin = lcl, ymax = ucl), alpha = 0.2)+
  facet_wrap(~label)
```


